---
title: "Loading scRNA-seq Data"
author: "Yoshinori Hayakawa"
output: 
  html_document:
    toc: true
    toc_depth: 4
    df_print: "paged"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading scRNA-seq Data

### Objective  

* Load expression matrix data and make it compatible with Seurat.  
* Add cell metadata to the Seurat object.  

### Data Loading

#### Load Required Libraries
```{r,message=FALSE}
# code:1-1
library(Seurat)
library(tidyverse)  # Includes dplyr, ggplot2, etc.
```

#### Load Data Information  
Load the path to the expression matrix data and a text file containing metadata.
```{r,message=FALSE}
# code:1-2
# Path to the text file summarizing expression matrix information
data_sheet_path <- "./data_info/datasheet.tsv"
```

```{r,message=FALSE}
# code:1-3
data_sheet_df <- read_tsv(data_sheet_path)
```

Check the contents of the data file.  
It contains metadata and paths to the datasets.
```{r,message=FALSE}
# code:1-4
data_sheet_df
```

Extract metadata associated with the expression data as vectors.
```{r,message=FALSE}
# Metadata vectors
# code:1-5
age_vec <- data_sheet_df %>% pull(age)
area_vec <- data_sheet_df %>% pull(area)
condition_vec <- data_sheet_df %>% pull(condition)
sample_vec <- paste0(age_vec, "_", area_vec, "_", condition_vec)

names(age_vec) <- sample_vec
names(area_vec) <- sample_vec
names(condition_vec) <- sample_vec
names(sample_vec) <- sample_vec

# Path vectors for expression matrix data
# code:1-6
path_vec <- data_sheet_df %>% pull(path)
names(path_vec) <- sample_vec
```

#### Load Expression Matrices  

```{r,message=FALSE}
# Manage multiple expression matrices as a list.
# code:1-7
orig_mtx_list <- list()
for(smn in names(path_vec)){
  print(smn)
  orig_mtx_list[[smn]] <- Read10X_h5(path_vec[smn])
}
```

Save the raw expression matrix data as an R object.
```{r,message=FALSE}
# code:1-8
saveRDS(orig_mtx_list, "./output/01_load/obj/orig_mtx_list.RDS")
```

### Create Seurat Object  
Convert the loaded expression matrix into a Seurat object.  
`min.cells = 3`: Genes expressed in fewer than 3 cells are removed.  
```{r,message=FALSE}
# code:1-9
# Prepare a list to store Seurat objects
created_seurat_obj_list <- list()

# Create Seurat objects
for(snm in names(orig_mtx_list)) {  # Iterate over multiple files
  print(snm)
  created_seurat_obj_list[[snm]] <- CreateSeuratObject(orig_mtx_list[[snm]], min.cells = 3, project = snm)
}
```

Check the created Seurat objects.
```{r,message=FALSE}
# code:1-10
created_seurat_obj_list
```

Check the metadata of the Seurat objects.
```{r,message=FALSE}
# code:1-11
head(created_seurat_obj_list[[1]][[]])
head(created_seurat_obj_list[[2]][[]])
```

#### Add Metadata to Seurat Object  

```{r,message=FALSE}
# code:1-12
for(smn in names(created_seurat_obj_list)){
  print(smn)
  created_seurat_obj_list[[smn]][[]]["age"]        <- age_vec[smn]
  created_seurat_obj_list[[smn]][[]]["area"]        <- area_vec[smn]
  created_seurat_obj_list[[smn]][[]]["condition"]   <- condition_vec[smn]
  created_seurat_obj_list[[smn]][[]]["sample_name"] <- sample_vec[smn]
}
```

Verify that metadata has been added.
```{r,message=FALSE}
# code:1-13
head(created_seurat_obj_list[[1]][[]])
head(created_seurat_obj_list[[2]][[]])
```

#### Save Seurat Object Before Quality Control  
```{r,message=FALSE}
# code:1-14
saveRDS(created_seurat_obj_list, "./output/01_load/obj/created_seurat_obj_list.RDS")
